\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{geometry}
\usepackage{rotating}
\usepackage{listings}
\usepackage{xcolor} 
\usepackage{tabularx}
\usepackage{booktabs}
\setlength{\parindent}{0pt}
\lstset{
  backgroundcolor=\color{lightgray},
  basicstyle=\ttfamily\small,
  keywordstyle=\color{blue},
  commentstyle=\color{green!60!black},
  stringstyle=\color{red},
  breaklines=true,
  showstringspaces=false
}

\geometry{a4paper, margin=1in}

\title{STAT 385 Final Project}
\author{Larry Barrett, Daniel Mroz, Al Pakrosnis}
\date{\today}

\begin{document}

\maketitle

\section{Question Synthesis}
Coming into this project our team wanted to separate analysis into three primary sections; the first two would be 

\section{Question 1}
\textbf{Q1. How does}
\subsection{Q1 EDA}

\newpage 

\section{Question 2}
\textbf{Q1. How does}
\subsection{Q2 EDA}

\newpage

\section{Question 3}
\subsection{Intro}
\textbf{Which students are most likely to need additional academic support, and what factors best predict this need?}

\begin{itemize}
    \item \textit{Operational Definition}: A student "needs support" if they have at least one past class failure or have received extra school-provided educational support.
    \item \textit{Goal}: Develop predictive models to classify students who need support, and identify key predictors
\end{itemize}

\textbf{Statistical and Machine Learning Solutions:}

\begin{itemize}
    \item Logistic Regression: For interpretable, baseline classification.
    \item Naive Bayes: For a simple probabilistic approach.
    \item Random Forest: For robust, non-linear modeling and feature importance.
\end{itemize}


\subsection{Q3 EDA}

\subsection{Data Preparation Steps:}
\begin{itemize}
    \item Using the vertically merged dataset, created a binary outcome variable \texttt{needs_support}, 1 if any failures or school support, 0 otherwise.
    \item Converted categorical variables to factors
    \item Removed grade columns and identifies for modeling
\end{itemize}


\section{Concluding Remarks}


\end{document}